---
title: "Preregistration ESM analysis code"
author: "Anu Hiekkaranta & Koen Derks"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 5
    number_sections: true
    df_print: paged
---

In this document we will describe the planned analyses for the experiment. To illustrate these analyses we will use an example data set, which is to be replaced when data collection has ended.

```{r Setting document options, echo = F}

# First we set some global document options so as that messages from R are not shown in the output
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center", 
                      fig.width = 10)

```

```{r Setting up global r environment}

# Clear the workspace
rm(list=ls())

# Load the required packages
library(lme4)		        # For frequentist hierarchical modeling
library(lmerTest)	      # For testing effect in lme4
library(brms) 	        # For Bayesian hierarchical modeling
library(bayesplot)	    # For plotting the samples
library(viridis)
library(ggmcmc)
library(ggthemes)
library(ggridges)
library(gridExtra)
library(mcmcplots)

# Set a global seed to make the results reproducible
seed <- 120495

# Specify the algorithmic options for `lme4`
ctrl <- lme4::lmerControl(optimizer = "optimx", calc.derivs = FALSE,
                          optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE))

# Specify the algorithmic options for `brms`
warmup          <- 2000          # Number of samples to discard before saving
iterations      <- 5000          # Number of total samples
chains          <- 4             # Number of individual sampling chains
cores           <- 4             # Number of CPU cores to use
max_treedepth   <- 10            # Maximum allowed tree depth
adapt_delta     <- 0.999         # Target acceptance rate of samples

# Helper functions for data manipulation and plotting
.scaleNumericData <- function(x, center = TRUE, scale = TRUE) {
  if (nrow(x) == 0)
    return(x)
  idx <- sapply(x, is.numeric)
  x[, idx] <- scale(x[, idx, drop = FALSE], center, scale)
  attr(x, which = "scaled:center") <- NULL
  attr(x, which = "scaled:scale")  <- NULL
  return(x)
}

# Create a function to visualize the posterior distributions of the fixed effects
.plotPopulationEffectDensities <- function(fit, title, xmin, xmax, warmup){
  colorValues <- c("darkorange1", "gold", "khaki4", "turquoise4", "mediumpurple3")
  mtfit <- ggs(fit)
  sumFixed <- summary(fit)$fixed
  parameters <- paste0("b_", rownames(sumFixed))
  plotData <- filter(mtfit, Parameter %in% parameters, Iteration > warmup)
  plotData$Parameter <- factor(plotData$Parameter, levels = parameters)
  p <- ggplot(plotData,
              aes(x = value, fill = Parameter)) +
              geom_density(alpha = .5) +
              labs(title = title) +
              geom_segment(x = 0, xend = 0, y = 0, yend = Inf, colour = "red", linetype = "solid", size = 0.5) + 
              scale_x_continuous(name   = "Standardized value", limits = c(xmin, xmax)) + 
              scale_y_continuous(name = "Density") +
              geom_segment(x = sumFixed[1,3], xend = sumFixed[1,3], y = 0, yend = Inf, 
                           col = colorValues[1], linetype = 2) +
              geom_segment(x = sumFixed[1,4], xend = sumFixed[1,4], y = 0, yend = Inf, 
                           col = colorValues[1], linetype = 2) +
              geom_segment(x = sumFixed[2,3], xend = sumFixed[2,3], y = 0, yend = Inf, 
                           col = colorValues[2], linetype = 2) +
              geom_segment(x = sumFixed[2,4], xend = sumFixed[2,4], y = 0, yend = Inf, 
                           col = colorValues[2], linetype = 2) +
              geom_segment(x = sumFixed[3,3], xend = sumFixed[3,3], y = 0, yend = Inf, 
                           col = colorValues[3], linetype = 2) +
              geom_segment(x = sumFixed[3,4], xend = sumFixed[3,4], y = 0, yend = Inf, 
                           col = colorValues[3], linetype = 2) +
              geom_segment(x = sumFixed[4,3], xend = sumFixed[4,3], y = 0, yend = Inf, 
                           col = colorValues[4], linetype = 2) +
              geom_segment(x = sumFixed[4,4], xend = sumFixed[4,4], y = 0, yend = Inf, 
                           col = colorValues[4], linetype = 2) +
              geom_segment(x = sumFixed[5,3], xend = sumFixed[5,3], y = 0, yend = Inf, 
                           col = colorValues[5], linetype = 2) +
              geom_segment(x = sumFixed[5,4], xend = sumFixed[5,4], y = 0, yend = Inf, 
                           col = colorValues[5], linetype = 2) +
              geom_segment(x = xmin, xend = xmax, y = -Inf, yend = -Inf) +
              geom_segment(x = -Inf, xend = -Inf, y = 0, yend = 20) +
              scale_fill_manual(name  = 'Parameters', 
                               values = colorValues, 
                               labels = c(expression(gamma['00']), 
                                          expression(gamma['10']),  
                                          expression(gamma['20']),
                                          expression(gamma['30']),
                                          expression(gamma['40']))) +
            theme_minimal() +
      theme(panel.grid = element_blank(), axis.ticks = element_line())
  return(p)
}

# Create a function to visualize the non-informed and informed prior and posterior distributions
.plotInfluenceOfPriors <- function(coef, fit_f, fit_ni, fit_i, title, priors, xmin, xmax, ymax){
  
  lmeEst <- summary(fit1_f)$coefficients[coef, "Estimate"]
  sumFixedni <- summary(fit1_bni)$fixed
  sumFixedi <- summary(fit1_bi)$fixed
  
  posterior1 <- posterior_samples(fit1_bni, pars = paste0("b_", coef))
  posterior2 <- posterior_samples(fit1_bi, pars = paste0("b_", coef))
  posterior1.2 <- bind_rows("prior 1" = gather(posterior1), "prior 2" = gather(posterior2), .id = "id")
  
  colorValues <- c("skyblue", "seagreen")
  row <- which(rownames(sumFixedi) == coef)
  
  p <- ggplot(data = posterior1.2,
        mapping = aes(x = value, fill = id, colour = key, linetype = key, alpha = key)) +
        geom_segment(x = 0, xend = 0, y = 0, yend = Inf, colour = "red", 
                     linetype = "solid", alpha = 1, size = 0.5) + 
        geom_segment(x = lmeEst, xend = lmeEst, y = 0, yend = Inf, colour = "black", 
                     linetype = "dashed", alpha = 1, size = 0.5) +      
        annotate(geom = "text", x = lmeEst, y = ymax - 0.2, hjust = -0.1,
                 label = paste0("LME estimate: ", round(lmeEst, 3)),
                 col = "black", family = theme_get()$text[["family"]], size = theme_get()$text[["size"]]/3.5,
                 fontface="italic") +
        geom_density(size = 0.75) +
        scale_x_continuous(name = "Standardized value", limits = c(xmin, xmax)) +
        scale_y_continuous(name = "Density", limits = c(0, ymax), breaks = pretty(c(0, ymax))) +
        scale_fill_manual(name = "Densities", values = colorValues, labels = priors) +
        scale_colour_manual(name = 'Posterior/Prior', values = c("black","red"), labels = c("posterior", "prior"))+
        scale_linetype_manual(name ='Posterior/Prior', values = c("solid","dotted"), labels = c("posterior", "prior"))+
        scale_alpha_discrete(name = 'Posterior/Prior', range  = c(.7,.3), labels = c("posterior", "prior"))+
        labs(title = title,
             subtitle = "2 different densities of priors and posteriors and the LME estimate") +
        geom_segment(x = xmin, xend = xmax, y = -Inf, yend = -Inf, colour = "black", linetype = "solid", alpha = 1) +
        geom_segment(x = -Inf, xend = -Inf, y = 0, yend = ymax, colour = "black", linetype = "solid", alpha = 1) +
        geom_segment(x = sumFixedni[row,3], xend = sumFixedni[row,3], y = 0, yend = Inf, 
                     col = colorValues[1], linetype = "dashed", alpha = 1) +
        geom_segment(x = sumFixedni[row,4], xend = sumFixedni[row,4], y = 0, yend = Inf, 
                     col = colorValues[1], linetype = "dashed", alpha = 1) +
        geom_segment(x = sumFixedi[row,3], xend = sumFixedi[row,3], y = 0, yend = Inf, 
                     col = colorValues[2], linetype = "dashed", alpha = 1) +
        geom_segment(x = sumFixedi[row,4], xend = sumFixedi[row,4], y = 0, yend = Inf, 
                     col = colorValues[2], linetype = "dashed", alpha = 1) +
        theme_minimal() +
        theme(panel.grid = element_blank(), axis.ticks = element_line())
  return(p) 
}

# Create a function to visualize the posterior distribution of the random slopes standard deviation
.plotPosteriorModelSD <- function(fit, var, title, xmin, xmax, color){
  parName <- paste0("sd_ID__", var)
  sumRandom <- summary(fit)$random$ID
  p <- ggplot(filter(ggs(fit), Parameter == parName, Iteration > warmup), aes(x = value)) +
      geom_segment(x = 0, xend = 0, y = Inf, yend = 0, col = "red") +
      geom_segment(x = sumRandom[2,1], xend = sumRandom[2,1], 
                   y = Inf, yend = 0, col = "blue") +
      geom_segment(x = sumRandom[2,3], xend = sumRandom[2,3], 
                   y = Inf, yend = 0, col = "blue", linetype = 2) +
      geom_segment(x = sumRandom[2,4], xend = sumRandom[2,4], 
                         y = Inf, yend = 0, col = "blue", linetype = 2) +
      geom_density(fill  = color, alpha = .05) +
      scale_x_continuous(name = "Standardized value", limits = c(xmin, xmax), breaks = pretty(c(xmin, xmax, 0))) + 
      scale_y_continuous(name = "Density") +
      labs(title = title) +
      geom_segment(x = xmin, xend = xmax, y = -Inf, yend = -Inf) +
      geom_segment(x = -Inf, xend = -Inf, y = 0, yend = 20) +
      theme_minimal() +
      theme(panel.grid = element_blank(), axis.ticks = element_line())
  return(p)
}

# Create a function to visualize the individual posterior distributions of the random slopes
.plotIndividualDifferences <- function(independent, dependent, fit, title, subtitle, xmin, xmax){
 
  samples <- posterior_samples(fit, pars = paste0(independent, "|ID"))
  samples <- as_tibble(t(samples[, 7:9]))
  
  col <- which(colnames(dataset) == dependent)
  dep <- aggregate(dataset[, col], by = list(ID = dataset$ID), FUN = mean, na.rm = T)
  colnames(dep) <- c("ID", dependent)
  dep <- dep[complete.cases(dep), ]
  
  plotdata <- cbind(dep, samples) %>%
  gather(key = "key", value = "value", -as.character(dependent), -ID) %>%
    group_by(ID) %>%
    mutate(meanperid = mean(value))%>%
    ungroup()
  
  plotdata$ID <- factor(plotdata$ID, levels = levels(plotdata$ID))
  
  p <- ggplot() +
        ggridges::geom_density_ridges(data  = plotdata, 
                                      aes(x = value, y = reorder(as.factor(ID), meanperid), 
                                          height = ..density.., fill = ID, scale = 3, alpha = .6),
                                          scale = 3, alpha = .6) +
        geom_point(data = summarise(group_by(plotdata, ID), meand = mean(meanperid)),
                    aes(x = meand, y = ID), col = "red") +
        scale_x_continuous(name = "Value", limits = c(xmin, xmax), breaks = pretty(c(xmin, xmax))) + 
        scale_y_discrete(name = "ID") +
        viridis::scale_fill_viridis(discrete = TRUE) +
        labs(title = title,
             subtitle = subtitle) +
        annotate(geom = "text", x = 1.5, y = 0.75, label = "Grand mean", col = "red", 
                 family = theme_get()$text[["family"]], size = theme_get()$text[["size"]]/3, fontface = "italic") +
        geom_vline(xintercept = 0, col = "red") +
        theme_minimal() +
        theme(panel.grid = element_blank(), axis.ticks = element_line(), axis.line.x = element_line())
  return(p)
}

# Apply the chosen seed globally
set.seed(seed)

```

# Data

The data set contains the following variables:

- $ID$: The identification number of the participant,
- $INEG$: The intensity of daily negative events experienced by the participant,
- $IPOS$: The intensity of daily positive events experienced by the participant,
- $RUM$: The intensity of ruminating experienced by the participant,
- $SAV$: The intensity of savoring experienced by the participant,
- $SHAREN$: The intensity by which the participant shared negative events,
- $SHAREP$: The intensity by which the participant shared positive events,
- $PP$: The psychopathology score of the participant,
- $GEN$: The gender of the participant,
- $AGE$: The age of the participant.

```{r Reading data set}

# Start by reading in the data set
fulldata <- read.csv(file = "https://raw.githubusercontent.com/koenderks/preregistrationESM/master/dataset.csv", sep = ",", fileEncoding="UTF-8-BOM")

# Select only the columns that are required for these analyses
fulldata <- fulldata[, c("ID", "INEG", "IPOS", "RUM", "SAV", "SHAREN", "SHAREP", "PP", "GEN", "AGE", "DAY", "C19SE", "C19SS")]

# Make factor variables for ID, DAY, and GEN
fulldata$ID   <- as.factor(fulldata$ID)
fulldata$DAY  <- as.factor(fulldata$DAY)
fulldata$GEN  <- as.factor(fulldata$GEN)

# Standardize the other numeric variables
fulldata <- .scaleNumericData(fulldata)

# Inspect the data set
head(fulldata)

```

# Hypotheses

The expected relationships in the data are the following:

- Expectation 1:  Sharing negative daily events is positively associated with psychopathology ($PP_{\uparrow} \rightarrow SHAREN_{\uparrow}$),
- Expectation 2:  Ruminating about daily negative events is positively associated with psychopathology ($PP_{\uparrow} \rightarrow RUM_{\uparrow}$),
- Expectation 3:  Sharing daily positive events is not associated with psychopathology ($PP_{\uparrow} \rightarrow SHAREP_{-}$),
- Expectation 4:  Savouring daily positive events is not associated with psychopathology ($PP_{\uparrow} \rightarrow SAV_{-}$),
- Expectation 5:  Intensity of daily negative events is positively associated with rumination ($INEG_{\uparrow} \rightarrow RUM_{\uparrow}$),
- Expectation 6:  Intensity of daily negative events is positively associated with sharing ($INEG_{\uparrow} \rightarrow SHAREN_{\uparrow}$),
- Expectation 7:  Intensity of daily positive events is positively associated with savouring ($IPOS_{\uparrow} \rightarrow SAV_{\uparrow}$),
- Expectation 8:  Intensity of daily positive events is positively associated with sharing ($IPOS_{\uparrow} \rightarrow SHAREP_{\uparrow}$),
- Expectation 9:  Psychopathology is positively associated with the intensity of daily negative events ($PP_{\uparrow} \rightarrow INEG_{\uparrow}$),
- Expectation 10: Psychopathology is not associated with the intensity of daily positive events ($PP_{\uparrow} \rightarrow IPOS_{-}$).

# Confirmatory Analyses

Below we will outline six models that will be use to find out to what degree the data contain support for any of the specified expectations.

## Model 1: Sharing negative events

Model 1 investigates the following expectations:

- Expectation 1: Talking more about daily negative events is associated with lower psychopathology,
- Expectation 6: Intensity of daily negative events is positively associated with sharing.

The regression formula for this model is:

$$SHAREN_{ij} = \gamma_{00} + \gamma_{10} \cdot PP_{ij} + \gamma_{20} \cdot AGE_{ij} + \gamma_{30} \cdot GEN_{{M}_{ij}} + \gamma_{40} \cdot INEG_{ij} + u_{4j} \cdot INEG_{ij} + u_{0j} + \epsilon_{ij},$$

in which observation $i$ from participant $j$ is predicted by a population-level intercept and coefficients for $PP$, $AGE$, $GEN_M$ and $INEG$ and a person-level intercept and random coefficient for $INEG$.

Therefore, to investigate expectation 1 and expectation 6, we test the following hypotheses respectively:

- Hypothesis 1: $H_0: \gamma_{10} \geq 0$ vs. $H_1: \gamma_{10} < 0$,
- Hypothesis 6: $H_0: \gamma_{40} \leq 0$ vs. $H_1: \gamma_{40} > 0$.

```{r class.source = "fold-show"}

# Hypothesis 1: 
# Talking more about daily negative events is associated with lower psychopathology 
# -> Negative population effect (PP < 0).
#
# Hypothesis 6: 
# Intensity of daily negative events is positively associated with sharing 
# -> Positive population effect (INEG > 0).
# 
# Model:
# SHAREN ~ 1 + PP + AGE + GEN + INEG + (1 + INEG|ID)

model1f <- SHAREN ~ 1 + PP + AGE + GEN + INEG + (1 + INEG|ID)
model1b <- SHAREN ~ 0 + Intercept + PP + AGE + GEN + INEG + (1 + INEG|ID)

```

```{r Preparing data for model 1}

# Remove NA values from the data set
dataset <- fulldata[!is.na(fulldata$INEG), ]

```

### Data visualization

First we visualize the data to get a better impression of the effects under investigation.

```{r Visualizing data for model 1}

p1 <- ggplot(data = dataset, aes(x = PP, y = SHAREN)) +
        geom_point(size = 1.2, alpha = 0.8, position = "jitter") +
        geom_smooth(method = lm, se = FALSE, size = .5, alpha = .8) +
        labs(title = "Linear Relationship for Sharing versus Intensity",
              subtitle = "The linear relationship is assumed to be equal across persons") +
        geom_segment(x = -3, xend = 3, y = -Inf, yend = -Inf, col = "black") +
        geom_segment(x = -Inf, xend = -Inf, y = -3, yend = 3, col = "black") +
        scale_y_continuous(name = "SHAREN", limits = c(-3, 3), breaks = pretty(c(-3, 3))) +
        scale_x_continuous(name = "PP", limits = c(-3, 3), breaks = pretty(c(-3, 3))) +
        theme_minimal() +
        theme(panel.grid = element_blank(), axis.ticks = element_line(), legend.position = "none")

p2 <- ggplot(data = dataset, aes(x = INEG, y = SHAREN, col = ID, group = ID)) +
        geom_point(size = 1.2, alpha = .8, position = "jitter") + 
        geom_smooth(method = lm, se = FALSE, size = .5, alpha = .8) +
        labs(title = "Linear Relationship for Different Persons as Observed",
              subtitle = "The linear relationship between the two is assumed to be different for all persons") +
        geom_segment(x = -3, xend = 3, y = -Inf, yend = -Inf, col = "black") +
        geom_segment(x = -Inf, xend = -Inf, y = -3, yend = 3, col = "black") +
        scale_y_continuous(name = "SHAREN", limits = c(-3, 3), breaks = pretty(c(-3, 3))) +
        scale_x_continuous(name = "INEG", limits = c(-3, 3), breaks = pretty(c(-3, 3))) +
        theme_minimal() +
        theme(panel.grid = element_blank(), axis.ticks = element_line())
grid.arrange(p1, p2, nrow = 1)

```

### Frequentist analysis

```{r Fitting model 1 using lme4}

fit1_f  <- lmerTest::lmer(formula = model1f, data = dataset, control = ctrl)
sum1_f  <- summary(fit1_f)
print(sum1_f)

``` 
##### ANOVA test for removal of random effects

By removing the random effect from the model, we can inspect whether it produced a better fit.

```{r Performing frequentist test for removal of random effects in model 1}

test1_f2 <- lmerTest::ranova(fit1_f)
print(test1_f2)

```

### Bayesian analysis

For the Bayesian analysis we use two different sets of priors.

First, we specify a set of non-informative priors on top of the `brms` default priors consisting of:

$$\gamma_{00} \sim N(0, 1) \\
  \gamma_{10} \sim N(0, 1) \\
  \gamma_{20} \sim N(0, 1) \\ 
  \gamma_{30} \sim N(0, 1) \\
  \gamma_{40} \sim N(0, 1)$$

Second, we specify a set of informed priors on top of the `brms` default priors consisting of: 

$$\gamma_{00} \sim Cauchy(-0.48, 0.85) \\
  \gamma_{10} \sim N(-.26, .25) \\
  \gamma_{20} \sim N(-0.01, 0.05) \\ 
  \gamma_{30} \sim N(0.45, .12) \\
  \gamma_{40} \sim N(0.48, 0.06)$$

```{r Fitting model 1 using brms}

# Non-informative priors:
#
# b_Intercept ~ Normal(0, 1)
# b_Age ~ Normal(0, 1)
# b_PP ~ Normal(0, 1)
# b_GEN2 ~ Normal(0, 1)
# b_INEG ~ Normal(0, 1)

priorsni <- c(set_prior("normal(0,1)", class = "b", coef = "Intercept"),
            set_prior("normal(0,1)", class = "b", coef = "AGE"),
            set_prior("normal(0,1)", class = "b", coef = "PP"),
            set_prior("normal(0,1)", class = "b", coef = "GEN2"),
            set_prior("normal(0,1)", class = "b", coef = "INEG"))

fit1_bni <- brms::brm(formula = model1b, 
                    data      = dataset, 
                    inits     = "random",
                    chains    = chains,
                    cores     = cores, 
                    prior     = priorsni,
                    warmup    = warmup, 
                    iter      = iterations,
                    seed      = seed,
                    sample_prior = TRUE,
                    control = list(adapt_delta = adapt_delta, 
                                   max_treedepth = max_treedepth),
                    save_pars = brms::save_pars(all = TRUE))

# Informed priors:
#
# b_Intercept ~ Cauchy(-0.48, 0.85)
# b_Age ~ Normal(-0.01, 0.05)
# b_PP ~ Normal(-.26,.25)
# b_GEN2 ~ Normal(0.45,.12)
# b_INEG ~ Normal(0.48,.06)

priorsi <- c(set_prior("cauchy(-.48,.85)", class = "b", coef = "Intercept"),
            set_prior("normal(-.01, .05)", class = "b", coef = "AGE"),
            set_prior("normal(-.26,.25)", class = "b", coef = "PP"),
            set_prior("normal(0.45,.12)", class = "b", coef = "GEN2"),
            set_prior("normal(0.48,.06)", class = "b", coef = "INEG"))

fit1_bi <- brms::brm(formula  = model1b, 
                     data     = dataset, 
                     inits    = "random",
                     chains   = chains,
                     cores    = cores, 
                     prior    = priorsi,
                     warmup   = warmup, 
                     iter     = iterations,
                     seed     = seed,
                     sample_prior = TRUE,
                     control = list(adapt_delta = adapt_delta, 
                                    max_treedepth = max_treedepth),
                     save_pars = brms::save_pars(all = TRUE))

```

#### Parameter estimates

##### Non-informative priors

```{r Printing summary of non-informed model 1}

sum1_bni <- summary(fit1_bni)
print(sum1_bni)

```

```{r Plotting chains of non-informed model 1}

mcmc_plot(fit1_bni, type = "trace")

```

```{r Plotting posterior histograms of non-informed model 1}

mcmc_plot(fit1_bni, type = "hist")

```

```{r Plotting posterior densities of coefficients of non-informed model 1}

.plotPopulationEffectDensities(fit1_bni, title = "Uninformed Posterior Density of Parameters With 95% CI lines", 
                                xmin = -1, xmax = 1, warmup = warmup)

```

##### Informed priors

```{r Printing summary of informed model 1}

sum1_bi <- summary(fit1_bi, priors = TRUE)
print(sum1_bi)

```

```{r Plotting chains of informed model 1}}

mcmc_plot(fit1_bi, type = "trace")

```

```{r Plotting posterior histograms of informed model 1}

mcmc_plot(fit1_bi, type = "hist")

```

```{r Plotting posterior densities of coefficients of informed model 1}

.plotPopulationEffectDensities(fit1_bi, title = "Informed Posterior Density of Parameters With 95% CI lines", 
                                xmin = -1, xmax = 1, warmup = warmup)

```

#### Hypothesis tests

##### Hypothesis 1: $\gamma_{10}$ < 0

```{r Compiling test table for fixed effect PP in model 1}

hypothesis1_bni <- brms::hypothesis(fit1_bni, "PP < 0", class = "b")
hypothesis1_bi <- brms::hypothesis(fit1_bi, "PP < 0", class = "b")
tab <- data.frame("Priors" = c("Non-informative", "Informed"))
tab <- cbind(tab, rbind(hypothesis1_bni$hypothesis, hypothesis1_bi$hypothesis))
print(tab)

```

```{r Plotting prior sensitivity for fixed effect PP in model 1}

.plotInfluenceOfPriors(coef = "PP", fit1_f, fit1_bni, fit1_bi, 
                    title = expression("Influence of (Informative) Priors on" ~ gamma[10]),
                    priors = c("Non-informative ~ N(0, 1) prior", "Informative ~ N(-.26,.25) prior"),
                    xmin = -2.5, xmax = 2.5, ymax = 6)

```

##### Hypothesis 6: $\gamma_{40}$ > 0

```{r Compiling test table for fixed effect INEG in model 1}

hypothesis6_bni <- brms::hypothesis(fit1_bni, "INEG > 0", class = "b")
hypothesis6_bi <- brms::hypothesis(fit1_bi, "INEG > 0", class = "b")
tab <- data.frame("Priors" = c("Non-informative", "Informed"))
tab <- cbind(tab, rbind(hypothesis6_bni$hypothesis, hypothesis6_bi$hypothesis))
print(tab)

```

```{r Plotting prior sensitivity for fixed effect INEG in model 1}

.plotInfluenceOfPriors(coef = "INEG", fit1_f, fit1_bni, fit1_bi, 
                    title = expression("Influence of (Informative) Priors on" ~ gamma[40]),
                    priors = c("Non-informative ~ N(0, 1) prior", "Non-informative ~ N(0.48, 0.06) prior"),
                    xmin = -2.5, xmax = 2.5, ymax = 7)

```

```{r Plotting posterior distributions of individual differences in model 1, fig.height = 10, fig.width = 15}

p1 <- .plotIndividualDifferences(independent = "INEG", dependent = "SHAREN", fit = fit1_bni, 
                           title = expression(paste("ID Level Deviation of Regression Coefficient of INEG on SHARIN (", u["4j"],")")),
                           subtitle = expression(paste("Uninformed Posterior distribution of deviation of regression coefficient (", u["4j"],") per ID")),
                           xmin = -1, xmax = 1)

p2 <- .plotIndividualDifferences(independent = "INEG", dependent = "SHAREN", fit = fit1_bi, 
                           title = expression(paste("ID Level Deviation of Regression Coefficient of INEG on SHARIN (", u["4j"],")")),
                           subtitle = expression(paste("Informed Posterior distribution of deviation of regression coefficient (", u["4j"],") per ID")),
                           xmin = -1, xmax = 1)

grid.arrange(p1, p2, nrow = 1)

```

## Model 2: Sharing positive events

```{r class.source = "fold-show"}

# Hypothesis 3: Talking more about daily positive events is not associated with psychopathology -> No population effect (PP = 0).
# Hypothesis 8: Intensity of daily positive events is positively associated with sharing -> Positive group effect (IPOS/ID > 0).

model2f <- SHAREN ~ 1 + PP + AGE + GEN + IPOS + (1 + IPOS|ID)
model2b <- SHAREN ~ 0 + Intercept + PP + AGE + GEN + IPOS + (1 + IPOS|ID)

```

## Model 3: Ruminating negative events

```{r class.source = "fold-show"}

# Hypothesis 2: Ruminating more about daily negative events is associated with higher psychopathology -> Positive population effect (PP > 0).
# Hypothesis 5: Intensity of daily negative events is positively associated with rumination. -> Positive group effect (INEG/DAY > 0).

model3f <- RUM ~ 1 + PP + AGE + GEN + INEG + (1 + INEG|ID)
model3b <- RUM ~ 0 + Intercept + PP + AGE + GEN + INEG + (1 + INEG|ID)

```

## Model 4: Savouring positive events

```{r class.source = "fold-show"}
# Hypothesis 4: Savouring daily positive events is not associated with psychopathology. -> No population effect (PP = 0)
# Hypothesis 7: Intensity of daily positive events is positively associated with savouring. -> Positive group effect (IPOS/DAY > 0)

model4f <- SAV ~ 1 + PP + AGE + GEN + IPOS + (1 + IPOS|ID)
model4b <- SAV ~ 0 + Intercept + PP + AGE + GEN + IPOS + (1 + IPOS|ID)
```

## Model 5: Intensity negative events

```{r class.source = "fold-show"}

# Hypothesis 9: Psychopathology is positively associated with the intensity of daily negative events. -> Positive population effect (PP > 0).

model5f <- INEG ~ 1 + PP + AGE + GEN
model5b <- INEG ~ 0 + Intercept + PP + AGE + GEN

```

## Model 6: Intensity positive events

```{r class.source = "fold-show"}

# Hypothesis 10: Psychopathology is not associated with the intensity of daily positive events. -> No population effect (PP = 0).

model6f <- IPOS ~ 1 + PP + AGE + GEN
model6b <- IPOS ~ 0 + Intercept + PP + AGE + GEN

```

# Confirmatory Results

```{r Creating results table}

types <- c("Frequentist", "Non-informed Bayesian", "Informed Bayesian")

# Model 1 hypothesis 1
model1_model <- c(1, NA, NA)
model1_hyp1_h0 <- c("PP >= 0", NA, NA)
model1_hyp1_h1 <- c("PP < 0", NA, NA)
model1_hyp1_estimates <- c(sum1_f$coefficients["PP", "Estimate"], sum1_bni$fixed["PP", "Estimate"], sum1_bi$fixed["PP", "Estimate"])
model1_hyp1_errors <- c(sum1_f$coefficients["PP", "Std. Error"], sum1_bni$fixed["PP", "Est.Error"], sum1_bi$fixed["PP", "Est.Error"])
model1_hyp1_p <- c(sum1_f$coefficients["PP", 5], NA, NA)
model1_hyp1_pp <- c(NA, hypothesis1_bni$hypothesis$Post.Prob, hypothesis1_bi$hypothesis$Post.Prob)

# Model 1 hypothesis 6
model1_hyp6_h0 <- c("INEG <= 0", NA, NA)
model1_hyp6_h1 <- c("INEG > 0", NA, NA)
model1_hyp6_estimates <- c(sum1_f$coefficients["INEG", "Estimate"], sum1_bni$fixed["INEG", "Estimate"], sum1_bi$fixed["INEG", "Estimate"])
model1_hyp6_errors <- c(sum1_f$coefficients["INEG", "Std. Error"], sum1_bni$fixed["INEG", "Est.Error"], sum1_bi$fixed["INEG", "Est.Error"])
model1_hyp6_p <- c(sum1_f$coefficients["INEG", 5], NA, NA)
model1_hyp6_pp <- c(NA, hypothesis6_bni$hypothesis$Post.Prob, hypothesis6_bi$hypothesis$Post.Prob)

tab <- data.frame(model = c(model1_model, model1_model),
                  h0 = c(model1_hyp1_h0, model1_hyp6_h0),
                  h1 = c(model1_hyp1_h1, model1_hyp6_h1),
                  type = rep(types, 2),
                  estimate = c(model1_hyp1_estimates, model1_hyp6_estimates),
                  error = c(model1_hyp1_errors, model1_hyp6_errors),
                  p = c(model1_hyp1_p, model1_hyp6_p),
                  postprob = c(model1_hyp1_pp, model1_hyp6_pp))
colnames(tab) <- c("Model", "H0", "H1", "Analysis type", "Estimate", "Est.Error", "p(y|H0)", "p(H1|y)")
print(tab)

```

# Exploratory Analyses

# Exploratory Results